---
title: "CS6140 Project"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
```

## Authors
- Abhay Kasturia [MSCS, fourth semester]
- Nakul Cammasamudram [MSCS, third semester]
- Philip Parker [MSCS, third semester]


## Data Cleaning and Exploration
```{r}
rm(list=ls())
# setwd("~/Workspace/northeastern/cs6140/CS6140_project/")
boston_cleaned <- read.csv("./data/boston_data.csv", sep = ",", header = TRUE)
head(boston_cleaned)
```

```{r}
listings <- read.csv("./data/listings.csv", sep = ",", header = TRUE)
head(listings)
colnames(listings)
```

### Features to delete 

```{r}
features_to_delete = c("id", "listing_url", "scrape_id", "last_scraped", "name", "summary", "space", "description", "experiences_offered", "neighborhood_overview", "notes", "transit", "access", "interaction", "house_rules", "thumbnail_url", "medium_url", "picture_url", "xl_picture_url", "host_id", "host_url", "host_name", "host_since", "host_location", "host_about", "host_response_time", "host_response_rate", "host_acceptance_rate", "host_thumbnail_url", "host_picture_url", "host_verifications", "host_has_profile_pic", "street", "neighbourhood_group_cleansed", "is_location_exact", "amenities", "calendar_updated", "has_availability", "availability_30", "availability_60", "availability_90", "availability_365", "calendar_last_scraped", "first_review", "last_review", "requires_license", "license", "jurisdiction_names", "require_guest_profile_picture", "require_guest_phone_verification", "host_neighbourhood", "neighbourhood", "neighbourhood_cleansed", "city", "state", "market", "smart_location", "country_code", "country", "latitude", "longitude", "maximum_nights", "review_scores_accuracy", "review_scores_cleanliness", "review_scores_checkin", "review_scores_communication", "review_scores_location", "review_scores_value", "square_feet", "weekly_price", "monthly_price", "security_deposit", "cleaning_fee", "extra_people", "host_total_listings_count", "calculated_host_listings_count", "host_listings_count")
```


### Filtered data
```{r}
boston <- listings %>% select(-one_of(features_to_delete))
head(boston)
colnames(boston)
summary(boston)
```

### Exploration
```{r}
summary(boston)
```

### Price distribution
```{r}

```


## Description of problem


## Split training and validation data
```{r}
ratio <- sample(1:nrow(boston), 0.7*nrow(boston))
boston_training <- boston[ratio, ]
boston_validation <- boston[-ratio, ]
```

## Summary of data
http://insideairbnb.com/get-the-data.html

```{r}
```

### One-variable summary statistics
```{r}
summary(boston_training)
```


### Two-variable summary statistics
```{r}
round(cor(boston_training), digits = 2)
pairs(boston_training)
```

### Findings
#### Highly correlated predictors

#### Categorical predictors

#### Missing values
```{r}
sum(is.na(training))
```


#### Outliers
 

## Methods


## Preliminary results


## References



