---
title: "CS6140_final_report"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(ggplot2)
library(dummies)
library(gmapsdistance)
library(hashmap)
library(gbm)
library(ggmap)
library(mgcv)
library(glmnet)
library(tidyr)
library(leaps)

set.seed(123)
```

# Import Data
```{r}
# Choose whether to reprocess data
reprocess = FALSE
reprocess = reprocess || !file.exists("../data/boston_data_raw.csv")
reprocess = reprocess || !file.exists("../data/boston_data.csv")
reprocess = reprocess || !file.exists("../data/boston_data_dummied.csv")

reprocess = reprocess || !file.exists("../data/boston_ddowntown.csv")
reprocess = reprocess || !file.exists("../data/boston_dairport.csv")
reprocess = reprocess || !file.exists("../data/boston_dboth.csv")

reprocess = reprocess || !file.exists("../data/boston_ddowntown_dummied.csv")
reprocess = reprocess || !file.exists("../data/boston_dairport_dummied.csv")
reprocess = reprocess || !file.exists("../data/boston_dboth_dummied.csv")

reprocess = reprocess || !file.exists("../data/boston_outliers.csv")

if (!reprocess) {
  # Read in existing data from file
  boston.data.raw <- read.csv("../data/boston_data_raw.csv", sep = ",", header=TRUE, na.strings=c("", " ", "NA"))
  boston.data <- read.csv("../data/boston_data.csv", sep = ",", header=TRUE, na.strings=c("", " ", "NA"))
  boston.dummied <- read.csv("../data/boston_data_dummied.csv", sep = ",", header=TRUE, na.strings=c("", " ", "NA"))
  
  boston.dboth <- read.csv("../data/boston_dboth.csv", sep = ",", header=TRUE, na.strings=c("", " ", "NA"))
  boston.ddowntown<- read.csv("../data/boston_ddowntown.csv", sep = ",", header=TRUE, na.strings=c("", " ", "NA"))
  boston.dairport <- read.csv("../data/boston_dairport.csv", sep = ",", header=TRUE, na.strings=c("", " ", "NA"))
  
  boston.dboth.dummied <- read.csv("../data/boston_dboth_dummied.csv", sep = ",", header=TRUE, na.strings=c("", " ", "NA"))
  boston.ddowntown.dummied <- read.csv("../data/boston_ddowntown_dummied.csv", sep = ",", header=TRUE, na.strings=c("", " ", "NA"))
  boston.dairport.dummeied <- read.csv("../data/boston_dairport_dummied.csv", sep = ",", header=TRUE, na.strings=c("", " ", "NA"))
  
  boston.outliers <- read.csv("../data/boston_outliers.csv", sep = ",", header=TRUE, na.strings=c("", " ", "NA"))
  
  full.data <- read.csv("../data/listings.csv", sep = ",", header=TRUE, na.strings=c("", " ", "NA"))
  
} else {
  # Read in full dataset
  full.data <- read.csv("../data/listings.csv", sep = ",", header=TRUE, na.strings=c("", " ", "NA"))
  
  
  # Select features to keep
  features_to_keep <- c("host_is_superhost", "host_identity_verified", "neighbourhood_cleansed", "property_type", "room_type", "accommodates", "bathrooms", "bedrooms", "beds", "bed_type", "price", "guests_included", "minimum_nights", "number_of_reviews", "instant_bookable", "is_business_travel_ready", "cancellation_policy")
  boston.data.raw <- full.data[ , features_to_keep, drop=FALSE]
  
  
  # Clean dataframe
  ## Omit NA values
  boston.data <- na.omit(boston.data.raw)
  
  ## Change price to numeric
  boston.data$price <- as.numeric(gsub(",", "", substr(boston.data$price, 2, length(boston.data$price) - 1)))
  
  ## Rename neighbourhood_cleansed to neighborhood
  names(boston.data)[names(boston.data) == "neighbourhood_cleansed"] <- "neighborhood"
  
  ## Keep up to 95th percentile of price
  value = quantile(boston.data$price, c(.95))[[1]]
  boston.outliers <- boston.data[boston.data$price > value, ]
  boston.data <- boston.data[boston.data$price <= value, ]
  
  
  
  # Dummy categorical features
  ## Remove categorical columns to be re-added
  categorical <- c("host_is_superhost", "host_identity_verified", "neighborhood", "property_type", "room_type", "bed_type", "instant_bookable", "is_business_travel_ready", "cancellation_policy")
  boston.dummied <- boston.data %>% select(-one_of(categorical))
  
  ## host_is_superhost
  boston.dummied <- cbind(boston.dummied, host_is_superhost=dummy(boston.data$host_is_superhost, sep="_")[ , -1])
  
  ## host_identity_verified
  boston.dummied <- cbind(boston.dummied, host_identity_verified=dummy(boston.data$host_identity_verified, sep="_")[ , -1])
  
  ## instant_bookable
  boston.dummied <- cbind(boston.dummied, instant_bookable=dummy(boston.data$instant_bookable, sep="_")[ , -1])
  
  ## is_business_travel_ready
  boston.dummied <- cbind(boston.dummied, is_business_travel_ready=dummy(boston.data$is_business_travel_ready, sep="_")[ , -1])
  
  ## property_type
  temp <- data.frame(dummy(boston.data$property_type))[ , -1]
  boston.dummied <- cbind(boston.dummied, temp)
  
  ## room_type
  temp <- data.frame(dummy(boston.data$room_type))[ , -1]
  boston.dummied <- cbind(boston.dummied, temp)
  
  ## bed_type
  temp <- data.frame(dummy(boston.data$bed_type))[ , -1]
  boston.dummied <- cbind(boston.dummied, temp)
  
  ## cancellation_policy
  temp <- data.frame(dummy(boston.data$cancellation_policy))[ , -1]
  boston.dummied <- cbind(boston.dummied, temp)
  
  
  # Construct distinct datasets

  ## Dataset with distance to downtown and airport
  boston.dboth.dummied <- boston.dummied
  boston.dboth.dummied$ddowntown <- 0
  boston.dboth.dummied$dairport <- 0
  
  boston.dboth <- boston.data
  boston.dboth$ddowntown <- 0
  boston.dboth$dairport <- 0
  
  ### Calculate driving distance from property location to downtown/airport
  ddowntown = hashmap(levels(boston.data$neighborhood), integer(length(levels(boston.data$neighborhood))))
  dairport = hashmap(levels(boston.data$neighborhood), integer(length(levels(boston.data$neighborhood))))

  for (i in 1:length(levels(boston.data$neighborhood))) {
    print(i)
    s <- levels(boston.data$neighborhood)[[i]]
    s2 <- paste(s, ", Boston MA")
    s2 <- gsub(" ", "+", s2, fixed=TRUE)
    ddowntown[[s]] <- gmapsdistance(origin=s2, destination="42.3555925+-71.0624982", mode="driving")[[2]]
    dairport[[s]] <- gmapsdistance(origin=s2, destination="42.3656171+-71.0117542", mode="driving")[[2]]
  }
  
  for (i in 1:nrow(boston.dboth.dummied)) {
    print(i)
    boston.dboth.dummied[i, "ddowntown"] <- ddowntown[[boston.data$neighborhood[[i]]]]
    boston.dboth.dummied[i, "dairport"] <- dairport[[boston.data$neighborhood[[i]]]]
    
    boston.dboth[i, "ddowntown"] <- ddowntown[[boston.data$neighborhood[[i]]]]
    boston.dboth[i, "dairport"] <- dairport[[boston.data$neighborhood[[i]]]]
  }
  
  ### Remove neighborhood columns
  boston.dboth.dummied <- boston.dboth.dummied[ , !(names(boston.dboth.dummied) %in% c("neighborhood"))]
  boston.dboth <- boston.dboth[ , !(names(boston.dboth) %in% c("neighborhood"))]
  
  ## Dataset with distance to downtown only
  boston.ddowntown.dummied <- boston.dboth.dummied[ , !(names(boston.dboth.dummied) %in% c("dairport"))]
  boston.ddowntown <- boston.dboth[ , !(names(boston.dboth) %in% c("dairport"))]
  
  ## Dataset with distance to airport only
  boston.dairport.dummied <- boston.dboth.dummied[ , !(names(boston.dboth.dummied) %in% c("ddowntown"))]
  boston.dairport <- boston.dboth[ , !(names(boston.dboth) %in% c("ddowntown"))]
  
  # Dummy neighborhood 
  ## neighborhood
  temp <- data.frame(dummy(boston.data$neighborhood))[ , -1]
  boston.dummied <- cbind(boston.dummied, temp)
  
  
  # Save data
  write.csv(boston.data.raw, file="../data/boston_data_raw.csv")
  write.csv(boston.data, file="../data/boston_data.csv")
  write.csv(boston.dummied, file="../data/boston_data_dummied.csv")
  
  write.csv(boston.ddowntown, file="../data/boston_ddowntown.csv")
  write.csv(boston.dairport, file="../data/boston_dairport.csv")
  write.csv(boston.dboth, file="../data/boston_dboth.csv")
  
  write.csv(boston.ddowntown.dummied, file="../data/boston_ddowntown_dummied.csv")
  write.csv(boston.dairport.dummied, file="../data/boston_dairport_dummied.csv")
  write.csv(boston.dboth.dummied, file="../data/boston_dboth_dummied.csv")
  
  write.csv(boston.outliers, file="../data/boston_outliers.csv")
}
```


# Preliminary Data Analysis
```{r, warning=FALSE}
# Visualizing the dataset on a map

# changing data for map visualization
full.data$price <- as.numeric(gsub(",", "", substr(full.data$price, 2, length(full.data$price) - 1)))
full.data <- full.data[!is.na(full.data$price),]
price_summary = summary(full.data$price)

# fetching data for map visualization
states <- map_data("state")
ma_df <- subset(states, region == "massachusetts")
counties <- map_data("county")
ma_county <- subset(counties, region == "massachusetts")
ma_base <- ggplot(data = ma_df, mapping = aes(x = long, y = lat, group = group)) + coord_fixed(1.3) +  geom_polygon(color = "black", fill = "gray")
ma_city<-ma_base + geom_polygon(data = ma_county, fill = NA, color = "white") +  geom_polygon(color = "black", fill = NA)  # get the state border back on top
ma_city + geom_point(aes(x=longitude,y=latitude,group=zipcode,color="red"),data=full.data)
map = get_map(location <- c(mean(full.data$longitude), mean(full.data$latitude)), zoom = 12, source = "google", maptype="roadmap")
ggmap(map)+ geom_point(aes(x=longitude,y=latitude,group=zipcode,color="red"),data=full.data)

# setting price categories
for (i in 1:nrow(full.data))
{
  if (full.data$price[i]<price_summary[2])
  {full.data$price_cat[i]=4}
  if (full.data$price[i]>price_summary[2]&full.data$price[i]<price_summary[4])
  {full.data$price_cat[i]=3}
  if (full.data$price[i]>price_summary[4]&full.data$price[i]<price_summary[5])
  {full.data$price_cat[i]=2}
  if (price_summary[4]<full.data$price[i])
  {full.data$price_cat[i]=1}
}

# plot map
ggmap(map) + geom_point(aes(x=longitude,y=latitude,group=price_cat,color=price_cat),data=full.data)
```


## Count Missing Values
```{r}
sum(is.na(boston.data.raw))
```


# Divide data into training and validation sets
```{r}
# Sample indices for training and validations sets
num_rows <- nrow(boston.data)

# 55% training
training <- sample(1:num_rows, floor(0.55 * num_rows))
rest <- (1:num_rows)[-training]

# 15% model selection
model_selection <- sample(rest, floor(0.15 * num_rows))
rest <- (1:num_rows)[-c(training, model_selection)]

# 15% validation #1
validation_1 <- sample(rest, floor(0.15 * num_rows))
rest <- (1:num_rows)[-c(training, model_selection, validation_1)]

# 15% validation #2
validation_2 <- sample(rest, floor(0.15 * num_rows))
rest <- (1:num_rows)[-c(training, model_selection, validation_1, validation_2)]

# Set datasets for each transformation
boston.data.training <- boston.data[training, ]
boston.data.model_selection <- boston.data[model_selection, ]
boston.data.validation_1 <- boston.data[validation_1, ]
boston.data.validation_2 <- boston.data[validation_2, ]
boston.data.model_selection.test <- boston.data.model_selection[ , "price"]
boston.data.validation_1.test <- boston.data.validation_1[ , "price"]
boston.data.validation_2.test <- boston.data.validation_2[ , "price"]

boston.dboth.training <- boston.dboth[training, ]
boston.dboth.model_selection <- boston.dboth[model_selection, ]
boston.dboth.validation_1 <- boston.dboth[validation_1, ]
boston.dboth.validation_2 <- boston.dboth[validation_2, ]
boston.dboth.model_selection.test <- boston.dboth.model_selection[ , "price"]
boston.dboth.validation_1.test <- boston.dboth.validation_1[ , "price"]
boston.dboth.validation_2.test <- boston.dboth.validation_2[ , "price"]

boston.ddowntown.training <- boston.ddowntown[training, ]
boston.ddowntown.model_selection <- boston.ddowntown[model_selection, ]
boston.ddowntown.validation_1 <- boston.ddowntown[validation_1, ]
boston.ddowntown.validation_2 <- boston.ddowntown[validation_2, ]
boston.ddowntown.model_selection.test <- boston.ddowntown.model_selection[ , "price"]
boston.ddowntown.validation_1.test <- boston.ddowntown.validation_1[ , "price"]
boston.ddowntown.validation_2.test <- boston.ddowntown.validation_2[ , "price"]

boston.dairport.training <- boston.dairport[training, ]
boston.dairport.model_selection <- boston.dairport[model_selection, ]
boston.dairport.validation_1 <- boston.dairport[validation_1, ]
boston.dairport.validation_2 <- boston.dairport[validation_2, ]
boston.dairport.model_selection.test <- boston.dairport.model_selection[ , "price"]
boston.dairport.validation_1.test <- boston.dairport.validation_1[ , "price"]
boston.dairport.validation_2.test <- boston.dairport.validation_2[ , "price"]
```


# Linear Regression

## Lasso Regression

```{r}

x.boston.data   <- model.matrix(price~.,boston.data.training)[,-5]
x.boston.dboth  <- model.matrix(price~.,boston.dboth.training)[,-5]
x.boston.ddowntown  <- model.matrix(price~.,boston.ddowntown.training)[,-5]
x.boston.dairport   <- model.matrix(price~.,boston.dairport.training)[,-5]

y.boston.data   <- boston.data.training$price
y.boston.dboth  <- boston.dboth.training$price
y.boston.ddowntown  <- boston.ddowntown.training$price
y.boston.dairport   <- boston.dairport.training$price

grid = 10^seq(15,-2, length = 100)
lasso.boston.data <- glmnet(x.boston.data,y.boston.data, alpha = 1, lambda = grid)
plot(lasso.boston.data, main = "Lasso regression \n Boston Data", label = TRUE, xvar = "lambda", xlim = c(-5,15))

lasso.boston.ddowntown <- glmnet(x.boston.ddowntown,y.boston.ddowntown, alpha = 1, lambda = grid)
plot(lasso.boston.ddowntown, main = "Lasso regression \n Boston Downtown Distance", label = TRUE, xvar = "lambda", xlim = c(-5,15))

lasso.boston.dairport <- glmnet(x.boston.dairport,y.boston.dairport, alpha = 1, lambda = grid)
plot(lasso.boston.dairport, main = "Lasso regression \n Boston Airport Distance", label = TRUE, xvar = "lambda", xlim = c(-5,15))

lasso.boston.dboth <- glmnet(x.boston.dboth,y.boston.dboth, alpha = 1, lambda = grid)
plot(lasso.boston.dboth, main = "Lasso regression \n Boston Both Distance", label = TRUE, xvar = "lambda", xlim = c(-5,15))

```

```{r}

cv.out.data <- cv.glmnet(x.boston.data,y.boston.data,alpha = 1)
plot(cv.out.data)

cv.out.ddowntown <- cv.glmnet(x.boston.ddowntown,y.boston.ddowntown,alpha = 1)
plot(cv.out.ddowntown)

cv.out.dairport <- cv.glmnet(x.boston.dairport,y.boston.dairport,alpha = 1)
plot(cv.out.dairport)

cv.out.dboth <- cv.glmnet(x.boston.dboth,y.boston.dboth,alpha = 1)
plot(cv.out.dboth)

```

```{r}

bestlam.lasso.data <- cv.out.data$lambda.min
cat("Best lambda Boston Data" , bestlam.lasso.data, "\n")
cat("Best log lambda Boston Data", log(bestlam.lasso.data), "\n")

bestlam.lasso.ddowntown <- cv.out.ddowntown$lambda.min
cat("Best lambda Boston Downtown" , bestlam.lasso.ddowntown, "\n")
cat("Best log lambda Boston Downtown" , log(bestlam.lasso.ddowntown), "\n")

bestlam.lasso.dairport <- cv.out.dairport$lambda.min
cat("Best lambda Boston Airport" , bestlam.lasso.dairport, "\n")
cat("Best log lambda Boston Airport" , log(bestlam.lasso.dairport), "\n")

bestlam.lasso.dboth <- cv.out.dboth$lambda.min
cat("Best lambda Boston Both" , bestlam.lasso.dboth, "\n")
cat("Best log lambda Boston Both" , log(bestlam.lasso.dboth), "\n")
```

```{r}
lasso.model.data <- glmnet(x.boston.data, y.boston.data, alpha=1, lambda = bestlam.lasso.data)
lasso.model.ddowntown <- glmnet(x.boston.ddowntown, y.boston.ddowntown, alpha=1, lambda = bestlam.lasso.ddowntown)
lasso.model.dairport <- glmnet(x.boston.dairport, y.boston.dairport, alpha=1, lambda = bestlam.lasso.dairport)
lasso.model.dboth <- glmnet(x.boston.dboth, y.boston.dboth, alpha=1, lambda = bestlam.lasso.dboth)

rmse <- function(test_data, model) {
  return(sqrt(mean((test_data$price - predict(model, newx = model.matrix(price~.,test_data)[,-5]))^ 2)))
}

cat("RMSE for Boston Data using Lasso", rmse(boston.data.validation_1, lasso.model.data), "$\n")
cat("RMSE for Boston Downtown Data using Lasso", rmse(boston.ddowntown.validation_1, lasso.model.ddowntown), "$\n")
cat("RMSE for Boston Airport Data using Lasso", rmse(boston.dairport.validation_1, lasso.model.dairport), "$\n")
cat("RMSE for Boston Both Data using Lasso", rmse(boston.dboth.validation_1, lasso.model.dboth), "$\n")
```

## Ridge Regression

```{r}
grid = 10^seq(15,-2, length = 100)
ridge.boston.data <- glmnet(x.boston.data,y.boston.data, alpha = 0, lambda = grid)
plot(ridge.boston.data, main = "Ridge regression \n Boston Data", label = TRUE, xvar = "lambda", xlim = c(-5,15))

ridge.boston.ddowntown <- glmnet(x.boston.ddowntown,y.boston.ddowntown, alpha = 0, lambda = grid)
plot(ridge.boston.ddowntown, main = "Ridge regression \n Boston Downtown Distance", label = TRUE, xvar = "lambda", xlim = c(-5,15))

ridge.boston.dairport <- glmnet(x.boston.dairport,y.boston.dairport, alpha = 0, lambda = grid)
plot(ridge.boston.dairport, main = "Ridge regression \n Boston Airport Distance", label = TRUE, xvar = "lambda", xlim = c(-5,15))

ridge.boston.dboth <- glmnet(x.boston.dboth,y.boston.dboth, alpha = 0, lambda = grid)
plot(ridge.boston.dboth, main = "Ridge regression \n Boston Both Distance", label = TRUE, xvar = "lambda", xlim = c(-5,15))
```

```{r}
cv.out.data <- cv.glmnet(x.boston.data,y.boston.data,alpha = 0)
plot(cv.out.data)

cv.out.ddowntown <- cv.glmnet(x.boston.ddowntown,y.boston.ddowntown,alpha = 0)
plot(cv.out.ddowntown)

cv.out.dairport <- cv.glmnet(x.boston.dairport,y.boston.dairport,alpha = 0)
plot(cv.out.dairport)

cv.out.dboth <- cv.glmnet(x.boston.dboth,y.boston.dboth,alpha = 0)
plot(cv.out.dboth)
```


```{r}
bestlam.ridge.data <- cv.out.data$lambda.min
cat("Best lambda Boston Data" , bestlam.ridge.data, "\n")
cat("Best log lambda Boston Data", log(bestlam.ridge.data), "\n")

bestlam.ridge.ddowntown <- cv.out.ddowntown$lambda.min
cat("Best lambda Boston Downtown" , bestlam.ridge.ddowntown, "\n")
cat("Best log lambda Boston Downtown" , log(bestlam.ridge.ddowntown), "\n")

bestlam.ridge.dairport <- cv.out.dairport$lambda.min
cat("Best lambda Boston Airport" , bestlam.ridge.dairport, "\n")
cat("Best log lambda Boston Airport" , log(bestlam.ridge.dairport), "\n")

bestlam.ridge.dboth <- cv.out.dboth$lambda.min
cat("Best lambda Boston Both" , bestlam.ridge.dboth, "\n")
cat("Best log lambda Boston Both" , log(bestlam.ridge.dboth), "\n")
```

```{r}
ridge.model.data <- glmnet(x.boston.data, y.boston.data, alpha=1, lambda = bestlam.ridge.data)
ridge.model.ddowntown <- glmnet(x.boston.ddowntown, y.boston.ddowntown, alpha=1, lambda = bestlam.ridge.ddowntown)
ridge.model.dairport <- glmnet(x.boston.dairport, y.boston.dairport, alpha=1, lambda = bestlam.ridge.dairport)
ridge.model.dboth <- glmnet(x.boston.dboth, y.boston.dboth, alpha=1, lambda = bestlam.ridge.dboth)

rmse <- function(test_data, model) {
  return(sqrt(mean((test_data$price - predict(model, newx = model.matrix(price~.,test_data)[,-5]))^ 2)))
}

cat("MSE for Boston Data using Ridge", rmse(boston.data.validation_1, ridge.model.data), "$\n")
cat("MSE for Boston Downtown Data using Ridge", rmse(boston.ddowntown.validation_1, ridge.model.ddowntown), "$\n")
cat("MSE for Boston Airport Data using Ridge", rmse(boston.dairport.validation_1, ridge.model.dairport), "$\n")
cat("MSE for Boston Both Data using Ridge", rmse(boston.dboth.validation_1, ridge.model.dboth), "$\n")
```

# GAM

```{r}
# Helper function

## @param test_data: data frame representing the test dataset
## @param model: GAM to be evaluated
## @returns: the root mean square error of the given GAM's predictions
rmse <- function(test_data, model) {
  return(sqrt(mean((test_data$price - predict.gam(model, test_data)) ^ 2)))
}
```

## Variable selection methods

Exploring three variable selection methods on the original dataset.

### Regression subset selction
```{r}
reg1 <- regsubsets(price~., data = boston.data.training, really.big = T)
reg1.summary <- summary(reg1)

par(mfrow = c(2,2))
plot(reg1.summary$rss, xlab = "Number of variables", ylab = "Residual Sum of Squares (RSS)", type = "l")
plot(reg1.summary$adjr2, xlab = "Number of variables", ylab = "Adjacent R square", type = "l")
plot(reg1.summary$cp, xlab = "Number of variables", ylab = "CP", type = "l")
plot(reg1.summary$bic, xlab = "Number of variables", ylab = "BIC", type = "l")

which.min(reg1.summary$bic)

# Results: neighborhood, room_type, bedrooms
```

```{r}
# GAM based on the above predictors
gam.var1 <- gam(price ~ neighborhood + room_type +s(bedrooms, k=5, bs="cr"), data=boston.data.training, method="REML")
summary(gam.var1)
# R^2: 0.563

gam.var1.rmse <- rmse(boston.data.model_selection, gam.var1)
gam.var1.rmse # 57.4993
```

### Forward Selection
```{r}
null <- lm(price~1, data=boston.data.training)
full <- lm(price~., data=boston.data.training)

forward_selection <- step(null, scope=list(lower=null, upper=full), direction="forward")
# Results: price ~ room_type + accommodates + neighborhood + bedrooms + property_type + cancellation_policy + instant_bookable + guests_included + bathrooms + minimum_nights + number_of_reviews + is_business_travel_ready + host_identity_verified + host_is_superhost

# The same predictors were obtained when "backward" and "both" directions for steps selctions as well.

# GAM based on the above predictors
gam.var2 <- gam(price ~ room_type + accommodates + neighborhood + bedrooms + property_type + cancellation_policy + instant_bookable + s(guests_included, bs="cr") + s(bathrooms, bs="cr") + s(minimum_nights, bs="cr") + s(number_of_reviews, bs="cr") + is_business_travel_ready + host_identity_verified + host_is_superhost, data=boston.data.training, method="REML")
summary(gam.var2)
# R^2: 0.609

gam.var2.rmse <- rmse(boston.data.model_selection, gam.var2)
gam.var2.rmse # 58.11913
```


### Shrinkage method within GAM
```{r}
gam.var3 <- gam(price ~ host_is_superhost + host_identity_verified + neighborhood + property_type + room_type + s(accommodates, bs="cs") + s(bathrooms, bs="cs") + s(bedrooms, k=5, bs="cs") + s(beds, bs="cs") + bed_type + s(guests_included, bs="cs") + s(minimum_nights, bs="cs") + s(number_of_reviews, bs="cs") + instant_bookable + is_business_travel_ready + cancellation_policy, data=boston.data.training, method="REML")
summary(gam.var3)
# R^2: 0.610

gam.var3.rmse <- rmse(boston.data.model_selection, gam.var3)
gam.var3.rmse # 55.14769
```

Since, the shrinkage method of variable selection has the best RMSE, it will be adopted as the variable selection method henceforth.

## GAM on all the datasets

### Original dataset
```{r}
gam.neighborhood <- gam(price ~ host_is_superhost + host_identity_verified + neighborhood + property_type + room_type + s(accommodates, bs="cs") + s(bathrooms, bs="cs") + s(bedrooms, k=5, bs="cs") + s(beds, bs="cs") + bed_type + s(guests_included, bs="cs") + s(minimum_nights, bs="cs") + s(number_of_reviews, bs="cs") + instant_bookable + is_business_travel_ready + cancellation_policy, data=boston.data.training, method="REML")
summary(gam.neighborhood)
# R^2: 0.610

gam.neighborhood.rmse <- rmse(boston.data.model_selection, gam.neighborhood)
gam.neighborhood.rmse # 55.14769
```

### Distance to both airport and downtown
```{r}
gam.dboth <- gam(price ~ host_is_superhost + host_identity_verified + property_type + room_type + s(accommodates, bs="cs") + s(bathrooms, bs="cs") + s(bedrooms, k=5, bs="cs") + +s(beds, bs="cs") + bed_type + s(guests_included, bs="cs") + s(minimum_nights, bs="cs") + s(number_of_reviews, bs="cs") + instant_bookable + is_business_travel_ready + cancellation_policy + s(ddowntown, bs="cs") + s(dairport, bs="cs"), data=boston.dboth.training, method="REML")
summary(gam.dboth)
# R^2: 0.600

gam.dboth.rmspe <- rmse(boston.dboth.model_selection, gam.dboth)
gam.dboth.rmspe # 55.77279
```

### Distance to airport
```{r}
gam.dairport <- gam(price ~ host_is_superhost + host_identity_verified + property_type + room_type + s(accommodates, bs="cs") + s(bathrooms, bs="cs") + s(bedrooms, k=5, bs="cs") + bed_type + s(guests_included, bs="cs") + s(minimum_nights, bs="cs") + s(number_of_reviews, bs="cs") + instant_bookable + is_business_travel_ready + cancellation_policy + s(dairport, bs="cs"), data=boston.dairport.training, method="REML")
summary(gam.dairport)
# R^2: 0.596

gam.dairport.rmse <- rmse(boston.dairport.model_selection, gam.dairport)
gam.dairport.rmse # 56.23814
```

### Distance to downtown
```{r}
gam.ddowntown <- gam(price ~ host_is_superhost + host_identity_verified + property_type + room_type + s(accommodates, bs="cs") + s(bathrooms, bs="cs") + s(bedrooms, k=5, bs="cs") + bed_type + s(guests_included, bs="cs") + s(minimum_nights, bs="cs") + s(number_of_reviews, bs="cs") + instant_bookable + is_business_travel_ready + cancellation_policy + s(ddowntown, bs="cs"), data=boston.ddowntown.training, method="REML")
summary(gam.ddowntown)
# R^2: 0.585

gam.ddowntown.rmse <- rmse(boston.ddowntown.model_selection, gam.ddowntown)
gam.ddowntown.rmse # 58.05393
```

### Results of the best model on validation_1 datasets

```{r}
gam.best <- gam.neighborhood
gam.best.v1_rmse <- rmse(boston.data.validation_1, gam.best)
gam.best.v1_rmse # 55.43666

# Most important predictors:
# Parametric: neighborhood, property_type, room_type, instant_bookable
# Non-parametric: accommodates, bedrooms, guests_included
```

# Regression Trees
## Train each of the transformations
```{r}
boston.data.boost <- gbm(price ~ ., boston.data.training, distribution="gaussian", n.trees=10000, cv.folds=5)
boston.dboth.boost <- gbm(price ~ ., boston.dboth.training, distribution="gaussian", n.trees=10000, cv.folds=5)
boston.ddowntown.boost <- gbm(price ~ ., boston.ddowntown.training, distribution="gaussian", n.trees=10000, cv.folds=5)
boston.dairport.boost <- gbm(price ~ ., boston.dairport.training, distribution="gaussian", n.trees=10000, cv.folds=5)
```

## Results of neighborhoods
```{r}
gbm.perf(boston.data.boost)
summary(boston.data.boost)
par(mfrow=c(1, 3))
plot(boston.data.boost, i=paste(summary(boston.data.boost, plotit=FALSE)$var[[1]], "", sep=""))
plot(boston.data.boost, i=paste(summary(boston.data.boost, plotit=FALSE)$var[[2]], "", sep=""))
plot(boston.data.boost, i=paste(summary(boston.data.boost, plotit=FALSE)$var[[3]], "", sep=""))
```

```{r}
yhat <- predict(boston.data.boost, newdata=boston.data.validation_1)
sqrt(mean((yhat - boston.data.validation_1.test)^2))
```

## Results of dboth
```{r}
gbm.perf(boston.dboth.boost)
summary(boston.dboth.boost)
par(mfrow=c(1, 3))
plot(boston.dboth.boost, i=paste(summary(boston.dboth.boost, plotit=FALSE)$var[[1]], "", sep=""))
plot(boston.dboth.boost, i=paste(summary(boston.dboth.boost, plotit=FALSE)$var[[2]], "", sep=""))
plot(boston.dboth.boost, i=paste(summary(boston.dboth.boost, plotit=FALSE)$var[[3]], "", sep=""))
```

```{r}
yhat <- predict(boston.dboth.boost, newdata=boston.dboth.validation_1)
sqrt(mean((yhat - boston.dboth.validation_1.test)^2))
```

## Results of ddowntown
```{r}
gbm.perf(boston.ddowntown.boost)
summary(boston.ddowntown.boost)
par(mfrow=c(1, 3))
plot(boston.ddowntown.boost, i=paste(summary(boston.ddowntown.boost, plotit=FALSE)$var[[1]], "", sep=""))
plot(boston.ddowntown.boost, i=paste(summary(boston.ddowntown.boost, plotit=FALSE)$var[[2]], "", sep=""))
plot(boston.ddowntown.boost, i=paste(summary(boston.ddowntown.boost, plotit=FALSE)$var[[3]], "", sep=""))
```

```{r}
yhat <- predict(boston.ddowntown.boost, newdata=boston.ddowntown.validation_1)
sqrt(mean((yhat - boston.ddowntown.validation_1.test)^2))
```

## Results of dairport
```{r}
gbm.perf(boston.dairport.boost)
summary(boston.dairport.boost)
par(mfrow=c(1, 3))
plot(boston.dairport.boost, i=paste(summary(boston.dairport.boost, plotit=FALSE)$var[[1]], "", sep=""))
plot(boston.dairport.boost, i=paste(summary(boston.dairport.boost, plotit=FALSE)$var[[2]], "", sep=""))
plot(boston.dairport.boost, i=paste(summary(boston.dairport.boost, plotit=FALSE)$var[[3]], "", sep=""))
```

```{r}
yhat <- predict(boston.dairport.boost, newdata=boston.dairport.validation_1)
sqrt(mean((yhat - boston.dairport.validation_1.test)^2))
```


# Best Model
## Results of the overall best model(GAM - Original dataset) on validation_2 dataset:
```{r}
gam.best.v2_rmse <- rmse(boston.data.validation_2, gam.best)
gam.best.v2_rmse # 52.29939
```




















